Research Design 2:\\ Comparing Reinforcement Learning and Finite State Machine Agents in Real Time Strategy Games: Impact on Player Experience

### 2025-05-25 Paper/main.tex
Paper is done. While introduction is long, there is more info than i normally would put there as a requirement.

\begin{abstract}
	abstract
\end{abstract}

\begin{IEEEkeywords}
	Keywords
\end{IEEEkeywords}

@@

\subsection{Positioning and Research Onion}
This research addresses the gap in player experience found in the literature, building on the works of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019} on AlphaStar by providing a
better understanding on the role RL agents will play in the future of RTS games. As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm,
following a deductive and experimental approach, gathering both quantitative and qualitative data to measure player experience.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.45\textwidth]{Images/Research_Onion.PNG}
	\caption{Research Onion}
	\label{fig:research_onion}
\end{figure}

@@

Game AI plays a huge role in player experience and immersion, as they provide the challenge and unpredictability that makes games fun and engaging.
It has evolved significantly over the years, especially in RTS games. Early RTS titles, such as StarCraft, relied on Finite State Machines (FSMs) for their AI decision-making. These FSM-based approaches,
while simple and easy to implement, are deterministic and predictable, which can lead to repetitive and boring gameplay, and allow players to exploit the gameplay patterns of the AI \cite{noauthor_finite_2020}
\cite{jagdale_finite_2021}.

% While extensive research has been conducted on the topic, most have focused more on the pure performance of the RL agent, and/or its impact on player experience, never directly
% comparing it to traditional FSMs, such as the work done by Grech \cite{grech_creating_2023}, Berta, et al. \cite{bin_ramlan_implementation_2021}, and Zhasulanov \cite{zhasulanov_enhancing_2024}.
% This study aims to address the research gap by directly comparing RL Agents to FSMs, and evaluating their impact on player experience, with the goal of identifying if the computational
% and development cost of implementing RL agents is justified by the improvement in player experience.

More recently, RL has emerged as an alternative AI approach, taking advantage of advancements in ML and computer hardware. In games such as AlphaStar \cite{vinyals_grandmaster_2019}, RL agents were able to
demonstrate adaptive and human-like behaviour, providing a more challenging and engaging experience for players. Another paper that highlights this is the work done by Grech \cite{grech_creating_2023},
where he created multiple difficulties of AI opponents using RL, and found that players reported higher levels of enjoyment and immersion. A similar study is the one done by \cite{bin_ramlan_implementation_2021},
where they trained an RL agent to act as an opponent in a fighting game, with the agent being able to adapt to the player's skill level, and provide a more engaging experience, similar to the work done by \cite{vinyals_grandmaster_2019}

@@

Independent variables are variables that are manipulated by the researcher, and are mainly used to influence the dependent variables. Dependent variables are what happen as a result of the independent variables,
and are what the researcher is interested in measuring.

@@

The difference between academic and non academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
than non-academic literature, which can easily be biased or contain false information. Academic literature can also be more in-depth and detailed, due to
the high research standards and requirements of academic institutions, especially IEEE.

@@

AI opponents play a critical role in maintaining this balance, as they provide the challenge and difficulty that the player must overcome. Traditionally FSMs are used for this, as they are simple to implement
and easy to understand, and when done correctly, provide a good challenge to the player. However, given enough time, players can learn the patterns of the FSMs, and exploit them,
making the game feel boring and leading to them falling out of the \textit{flow state} \cite{noauthor_comparative_nodate}. It is possible to combat this through weakening the player, or making the AI more difficult,
as is done in Souls-like games, however this can prove too challenging and overwhelm the player, as well as making the percieved balance seem unfair, once again breaking the flow \cite{jagdale_finite_2021} \cite{noauthor_finite_2020}.

RL agents, on the other hand, are able to learn how to play the game, and as such, adapt to the player and given scenario. This creates a more engaging
and immersive experience, as the player feels like they are playing against a real opponent, and not just a computer. This is especially true in RTS games,
due to the complexity of the game, and the many strategies that can be taken, which can be seen in the works done by \cite{vinyals_grandmaster_2019} and \cite{grech_creating_2023}.
A common tool used in most of these papers is the Unity ML-Agents toolkit, which is a framework for training RL agents in Unity games, making it easier to implement and train RL agents for reasearch purposes.
Along with that, most used the Proximal Policy Optimization (PPO) algorithm, which is a popular RL algorithm that is used for training agents in continuous action spaces, which is especially useful in real-time games.

In the work done by \cite{vinyals_grandmaster_2019}, they trained an RL agent to play StarCraft II, and found that the agent was albe to not only learn how to play the game, but also adapt to the player's actions and strategies in real time.
A similar result was found in the works done by \cite{bin_ramlan_implementation_2021} and \cite{raut_unity_2024}, where both used the Unity ML-Agents toolkit with PPO to train RL agents in 2 different genres.
\cite{bin_ramlan_implementation_2021} trained an RL agent to play a fighting game, with rewards for moving closer to the player, landing attacks and winning, and penalties for being hit, missing and losing.
\cite{raut_unity_2024} trained an RL agent to play a racing game, with the reward structure being based on the distance travelled and the time taken, which punishes the agent for crashing and/or taking too long.
Combined, these 3 papers show the power of RL agents in real-time games, and their adaptability to the player, however they all share the same 2 flaws. The agents are computationally expensive to train, and to
run, requiring a lot of time and resources not only for the developers, but also the players as they need to have a powerful enough computer to run the game and agent simultaneously. Along with that, the agents
can become too difficult to play against, and as such break the flow state, as the player is unable to keep up with the agent's actions and strategies, especially if they are new to the game.

\cite{grech_creating_2023} tackled this issue by saving multiple snapshots of the agent during training, and using them to create different difficulties of AI opponents, similarly to how FSM diffifulties are created.
Their results showed that players reported higher levels of enjoyment and immersion when playing against the RL agents, just like the previous works, with the key difference that new players were also able to enjoy
the game. Their implementation was rather simple, and as such requires more research to be done in improving it and identifying if it is a viable solution to the issues of RL agents. Another option would be
to build in a dynamic difficulty adjustment (DDA) system into either the agent or the game, as suggested by \cite{grech_creating_2023}, which would keep the rl agent's difficulty in check, while keeping its
adaptability and unpredictability. This shows the gap in literature, both in the comparison of RL and FSM AI opponents, but also in managing the diffuculty of RL agents.

@@

\begin{itemize}
	\item Raut, U., Galchhaniya, P., Nehete, A., Shinde, R., \& Bhoite, A. (2024). \textit{Unity ML-Agents: Revolutionizing Gaming Through Reinforcement Learning}. \cite{raut_unity_2024}
	\item Bin Ramlan, A. A., Ali, A. M., Abdul Hamid, N. H., \& Osman, R. (2021). \textit{The Implementation of Reinforcement Learning Algorithm for AI Bot in Fighting Video Game}. \cite{bin_ramlan_implementation_2021}
	\item Liu, R.-Z., Pang, Z.-J., Meng, Z.-Y., Wang, W., Yu, Y., \& Lu, T. (2022). \textit{On Efficient Reinforcement Learning for Full-length Game of StarCraft II}. \cite{liu_efficient_2022}
	\item Abdulghani, A. M., Abdulghani, M. M., Walters, W. L., \& Abed, K. H. (2023). \textit{Multi-Agent Reinforcement Learning System Using Value-Decomposition Network Algorithm in StarCraft Environment}. \cite{abdulghani_multi-agent_2023}
	\item Vinyals, O., Babuschkin, I., Czarnecki, W. M., et al. (2019). \textit{Grandmaster level in StarCraft II using multi-agent reinforcement learning}. \cite{vinyals_grandmaster_2019}
	\item Berta, R., Lazzaroni, L., Capello, A., et al. (2024). \textit{Development of Deep-Learning-Based Autonomous Agents for Low-Speed Maneuvering in Unity}. \cite{berta_development_2024}
	\item Jagdale, D. (2021). \textit{Finite State Machine in Game Development}. \cite{jagdale_finite_2021}
	\item \textit{Comparative Analysis of Game Development Techniques: Using Finite State Machine, Physics Simulation, Path Finding, Event Handling}. \cite{noauthor_comparative_nodate}
\end{itemize}

@@

The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determin their impact on player experience and immersion,
and what are the key factors that cause this impact/infuence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
and then conducting a play test experiment with players, where both groups will then be surveyed to gather data on their experience. Along with this,
data gathered during the playtest through unity analytics will be used to measure player engagement and immersion.

@@

This study adopts a \textbf{positivist research philosophy}, which means it emphasizes the use of objective measurements and observable phenomena, and it is suitable for this study as it aims
to evaluate the impact of RL and FSM AI opponents on player experience through measurable data, aligning it with \textbf{quantitative methodology}. It follows a \textbf{deductive approach}, with an \textbf{experimental research design},
as it starts with a hypothesis, and then using an experimental prototype, tests the hypothesis by allowing participants to play it, and then uses the data to observe the difference the independent variable (AI type)
has on the dependent variables \textbf(player experience, immersion, and perceived difficulty). This will then be used to answer the research questions and prove the hypothesis. It is important to note that the study
will also gather a very small amount of qualitative data, making it a \textbf{mixed-methods approach}. This is done to help align the more human aspects of player experience with the more quantitative aspects.

@@

% Is more to be done?
Create a simple RTS game, and implement both RL and FSM agents. Conduct a deductive experiment, where players will be split into 2 groups,
where the first group will play against the FSM, fill out a questionnaire to gather data on their experience, and then play against the RL agent,
and again answer the questionnaire. The second group will do the same, but in reverse order, playing against the RL agent first, and then the FSM.
The end of each experiment's questionnaires will have an open ended question, where players can provide feedback on their experience,
which will be used to gather qualitative data on their experience. Qualitative data will be used to support the quantitative data gathered from
the questionnaires, as well as some in-game analytics and logs. This should help capture the more qualitative aspects of player experience,
while still keeping everything quantitative and measurable. This should help in answering the research questions, and in determining the validity
of the Hypothesis.

@@

\textbf{Validity}: The study ensures validity by desigining the experiment to control for confounding variables, such as the order in which participants play, since on the second playtest they would be more
familiar with the game mechanics and controls. The use of validating the survey results with in-game analytics and engagement metrics further streghtens the validity of the research being conducted.

\textbf{Reliability}: To help ensure reliability, the same setup, game environment and surveys will be used for all participants. Any and all instructions will be standardised, using text and/or video/audio
recordings to ensure that all participants are given the same instructions, without any bias that comes from the researcher. All datacollection processes will be automated, further standardising the process
and ensuring reliability.

@@

\subsection{Etical considerations}

@@

% TODO: Discuss the following:
% You could add a short paragraph explicitly discussing validity (e.g., sample size, playtest design), reliability (consistency of results), and generalizability (applicability to other games or genres). 
% You touch on these in your methodology, but a brief mention here would strengthen the critique.

\subsection{Summary of Findings}
From the analysis of the playtest data, we saw that on easy difficulty, FSM agents provided a more meaningful challenge, and a better player experience than RL agents.
This could have been due to players not having learned how to exploit the FSM agents yet, or the RL agent being too inconsitent at this difficulty.
As the difficulty increased, RL agents became more competitive, with the average score difference decreasing significantly, and player experience metrics such as engagement and immersion increasing.
Contrarily, FSMs became more predictable and easier to exploit, leading to a decrease in player engagement and immersion at higher difficulties.
This means that hypothesis is only partially supported, as while players reported higher levels of enjoyment and immersion when playing against RL agents, this was not the case at the easy difficulty.

@@

This study set out to compare the impact of Reinforcement Learning (RL) and Finite State Machine (FSM) agents on player experience in Real-Time Strategy (RTS) games.
By implementing both AI approaches in a custom RTS prototype and conducting a controlled playtest, we gathered quantitative and qualitative data on player performance, engagement,
and immersion across varying difficulty levels.

The results demonstrate that RL agents, particularly at higher difficulties, provide a more engaging and immersive experience for players compared to FSM agents.
While FSMs can offer a meaningful challenge at lower levels, their predictability becomes apparent as difficulty increases, leading to reduced player engagement. In contrast, RL agents adapt to player strategies, maintaining challenge and interest, but can sometimes become too difficult or unpredictable, which may cause frustration for some players.

These findings support the hypothesis that RL agents can enhance player experience in RTS games, justifying their higher development and computational costs in contexts where player engagement and
immersion are priorities. However, careful difficulty balancing is essential to prevent RL agents from overwhelming players, especially less experienced ones.

% TODO: Discuss the limitations of the study, such as the small sample size, and the fact that it was only conducted on a single game, and not multiple games, or real world games.

% TODO: Discuss the implementation of DDA to better close the gap between difficulties, and account for the unpredictability of the RL agent, especially at earlier difficulties where convergence on a strategy
% is not complete.

% AI suggestions after prompt

% \subsection{Recommendations}

% Based on the findings of this study, the following recommendations are proposed for game developers and researchers:

% \begin{itemize}
%     \item \textbf{Adopt RL agents for advanced AI in RTS games} where player engagement and adaptability are desired, especially at higher difficulty levels.
%     \item \textbf{Implement dynamic difficulty adjustment (DDA)} or similar balancing mechanisms to ensure RL agents remain challenging but fair, preventing player frustration.
%     \item \textbf{Consider hybrid approaches} that combine the predictability and accessibility of FSMs at lower difficulties with the adaptability of RL agents at higher levels,
% 	catering to a wider range of player skills.
%     \item \textbf{Further research is needed} to explore long-term player retention, the impact of RL agents in other game genres, and the effectiveness of different balancing strategies.
%     \item \textbf{Evaluate computational and development costs} in the context of the target audience and desired player experience, as RL agents may not be necessary for all games or player bases.
% \end{itemize}

% In summary, RL agents have significant potential to improve player experience in RTS games, but their implementation should be carefully managed to maximize benefits and minimize drawbacks.

>>>>

\begin{abstract}
	In this study, we investigate the impact of Reinforcement Learning (RL) agents compared to traditional Finite State Machines (FSM)s on player experience.
	A simple Real-Time Strategy (RTS) game was developed to test whether RL agents provide a better player experience than FSMs. After a playtest with 2 
	groups of participants, we found that RL agents generally provided a more engaging and immersive experience, especially at higher levels of difficulty,
	while FSMs were preferred at easier levels. The results indicate that the development time and computational cost of RL agents may be justified in RTS games.
\end{abstract}

\begin{IEEEkeywords}
	Reinforcement Learning (RL), Finite State Machine (FSM), Real-Time Strategy (RTS) Games, Player Experience, Game AI
\end{IEEEkeywords}

@@

\subsection{Positioning and Research Onion}
This research addresses the gap in player experience found in the literature, being that while a lot of studies have been conducted on the performance or RL agents,
very few focused on the player experience difference between them, and FSMs. 
% By building on the works of \cite{grech_creating_2023} and \cite{vinyals_grandmaster_2019} on AlphaStar, this study will address the gap by
% providing a better understanding on the role RL agents will play in the future of RTS games. 
As can be seen in Figure \ref{fig:research_onion}, this study will follow a positivist research paradigm,
following a deductive and experimental approach, gathering both quantitative and qualitative data to measure player experience.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.4\textwidth]{Images/Research_Onion.PNG}
	\caption{Research Onion}
	\label{fig:research_onion}
\end{figure}

@@

Game AI plays a huge role in player experience and immersion, as they provide the challenge and unpredictability that makes games fun and engaging.
It has evolved significantly over the years, especially in RTS games. Early RTS titles, such as StarCraft, relied on Finite State Machines (FSMs) for their AI decision-making \cite{vinyals_grandmaster_2019}. 
These FSM-based approaches, while simple and easy to implement, are deterministic and predictable, which can lead to repetitive and boring gameplay,
and allow players to exploit the gameplay patterns of the AI \cite{noauthor_finite_2020} \cite{jagdale_finite_2021}.

More recently, RL has emerged as an alternative AI approach, taking advantage of advancements in ML and computer hardware. In games such as AlphaStar \cite{vinyals_grandmaster_2019}, RL agents were able to
demonstrate adaptive and human-like behaviour, providing a more challenging and engaging experience for players. Another paper that highlights this is the work done by Grech \cite{grech_creating_2023},
where he created multiple difficulties of AI opponents using RL, and found that players reported higher levels of enjoyment and immersion. 
% Seems excessive for an introduction and background section, and most of these are already in the literature review.
% A similar study is the one done by \cite{bin_ramlan_implementation_2021},
% where they trained an RL agent to act as an opponent in a fighting game, with the agent being able to adapt to the player's skill level, and provide a more engaging experience, similar to the work done by \cite{vinyals_grandmaster_2019}

@@

Independent variables are variables that are manipulated by the researcher and are mainly used to influence the dependent variables. 
Dependent variables are what happen because of the independent variables and are what the researcher is interested in measuring.

@@

The difference between academic and non-academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
than non-academic literature, which can easily be biased or contain false information. Academic literature can also be more in-depth and detailed, due to
the high research standards and requirements of academic institutions, especially IEEE.

@@

AI opponents play a critical role in maintaining this balance, as they provide the challenge and difficulty that the player must overcome. Traditionally FSMs are used for this, as they are simple to implement
and easy to understand, and when done correctly, provide a good challenge to the player. However, given enough time, players can learn the patterns of the FSMs, and exploit them,
making the game feel boring and leading to them falling out of the \textit{flow state} \cite{noauthor_comparative_nodate}. It is possible to combat this through weakening the player, or making the AI more difficult,
as is done in Souls-like games, however this can prove too challenging and overwhelm the player, as well as making the perceived balance seem unfair, once again breaking the flow \cite{jagdale_finite_2021} \cite{noauthor_finite_2020}.

RL agents, on the other hand, are able to learn how to play the game, and as such, adapt to the player and given scenario. This creates a more engaging
and immersive experience, as the player feels like they are playing against a real opponent, and not just a computer. This is especially true in RTS games,
due to the complexity of the game, and the many strategies that can be taken, which can be seen in the works done by \cite{vinyals_grandmaster_2019} and \cite{grech_creating_2023}.
A common tool used in most of these papers is the Unity ML-Agents toolkit, which is a framework for training RL agents in Unity games, making it easier to implement and train RL agents for research purposes.
Along with that, most used the Proximal Policy Optimization (PPO) algorithm, which is a popular RL algorithm that is used for training agents in continuous action spaces, which is especially useful in real-time games.

In the work done by \cite{vinyals_grandmaster_2019}, they trained an RL agent to play StarCraft II, and found that the agent was able to not only learn how to play the game, but also adapt to the player's actions and strategies in real time.
A similar result was found in the works done by \cite{bin_ramlan_implementation_2021} and \cite{raut_unity_2024}, where both used the Unity ML-Agents toolkit with PPO to train RL agents in 2 different genres.
\cite{bin_ramlan_implementation_2021} trained an RL agent to play a fighting game, with rewards for moving closer to the player, landing attacks and winning, and penalties for being hit, missing and losing.
\cite{raut_unity_2024} trained an RL agent to play a racing game, with the reward structure being based on the distance travelled and the time taken, which punishes the agent for crashing and/or taking too long.
Combined, these 3 papers show the power of RL agents in real-time games, and their adaptability to the player, however they all share the same 2 flaws. The agents are computationally expensive to train, and to
run, requiring a lot of time and resources not only for the developers, but also the players as they need to have a powerful enough computer to run the game and agent simultaneously. Along with that, the agents
can become too difficult to play against, and as such break the flow state, as the player is unable to keep up with the agent's actions and strategies, especially if they are new to the game.

\cite{grech_creating_2023} tackled this issue by saving multiple snapshots of the agent during training and using them to create different difficulties of AI opponents, similarly to how FSM difficulties are created.
Their results showed that players reported higher levels of enjoyment and immersion when playing against the RL agents, just like the previous works, with the key difference that new players were also able to enjoy
the game. Their implementation was rather simple, and as such requires more research to be done in improving it and identifying if it is a viable solution to the issues of RL agents. Another option would be
to build in a dynamic difficulty adjustment (DDA) system into either the agent or the game, as suggested by \cite{grech_creating_2023}, which would keep the rl agent's difficulty in check, while keeping its
adaptability and unpredictability. This shows the gap in literature, both in the comparison of RL and FSM AI opponents, but also in managing the difficulty of RL agents.

@@

% Takes up too much space, and most are found in the references section anyway.
\begin{itemize}
	\item Grech G. (2023). \textit{Creating Difficulty Levels with Reinforcement Learning in a Strategy Game}. \cite{grech_creating_2023}
	\item Raut, U., Galchhaniya, P., Nehete, A., Shinde, R., \& Bhoite, A. (2024). \textit{Unity ML-Agents: Revolutionizing Gaming Through Reinforcement Learning}. \cite{raut_unity_2024}
	\item Bin Ramlan, A. A., Ali, A. M., Abdul Hamid, N. H., \& Osman, R. (2021). \textit{The Implementation of Reinforcement Learning Algorithm for AI Bot in Fighting Video Game}. \cite{bin_ramlan_implementation_2021}
	\item Liu, R.-Z., Pang, Z.-J., Meng, Z.-Y., Wang, W., Yu, Y., \& Lu, T. (2022). \textit{On Efficient Reinforcement Learning for Full-length Game of StarCraft II}. \cite{liu_efficient_2022}
	% \item Abdulghani, A. M., Abdulghani, M. M., Walters, W. L., \& Abed, K. H. (2023). \textit{Multi-Agent Reinforcement Learning System Using Value-Decomposition Network Algorithm in StarCraft Environment}. \cite{abdulghani_multi-agent_2023}
	\item Vinyals, O., Babuschkin, I., Czarnecki, W. M., et al. (2019). \textit{Grandmaster level in StarCraft II using multi-agent reinforcement learning}. \cite{vinyals_grandmaster_2019}
	% \item Berta, R., Lazzaroni, L., Capello, A., et al. (2024). \textit{Development of Deep-Learning-Based Autonomous Agents for Low-Speed Maneuvering in Unity}. \cite{berta_development_2024}
	\item Jagdale, D. (2021). \textit{Finite State Machine in Game Development}. \cite{jagdale_finite_2021}
	% \item \textit{Comparative Analysis of Game Development Techniques: Using Finite State Machine, Physics Simulation, Path Finding, Event Handling}. \cite{noauthor_comparative_nodate}
\end{itemize}

@@

The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determine their impact on player experience and immersion,
and what are the key factors that cause this impact/influence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
and then conducting a play test experiment with players, where both groups will then be surveyed to gather data on their experience. Along with this,
data gathered during the playtest through unity analytics will be used to measure player engagement and immersion.

@@

This study adopts a \textbf{positivist research philosophy}, which means it emphasizes the use of objective measurements and observable phenomena, and it is suitable for this study as it aims
to evaluate the impact of RL and FSM AI opponents on player experience through measurable data, aligning it with \textbf{quantitative methodology}. It follows a \textbf{deductive approach}, with an \textbf{experimental research design},
as it starts with a hypothesis, and then using an experimental prototype, tests the hypothesis by allowing participants to play it, and then uses the data to observe the difference the independent variable (AI type)
has on the dependent variables \textbf(player experience, immersion, and perceived difficulty). This will then be used to answer the research questions and prove the hypothesis. It is important to note that the study
will also gather an fair amount of qualitative data, making it a \textbf{mixed-methods approach}. This is done to help align the more human aspects of player experience with the more quantitative aspects.

@@

A simple RTS game was created, using both RL and FSM agents, and a deductive experiment was conducted. The players were split into 2 groups,
where the first group played against the FSM first, then RL, across all difficulties, and the second group did the same, but with RL first.
After each difficulty, they were asked to fill out a questionnaire with quantitative and open-ended questions about their experience.
The qualitative data will help capture the more human aspects of player experience, and along with in-game analytics and logs, 
will be used to validate the data gathered from the questionnaires. This will help in answering the research questions, and keep everything measurable and comparable.

@@

\textbf{Validity}: The study ensures validity by designing the experiment to control for confounding variables, such as the order in which participants play, since on the second playtest they would be more
familiar with the game mechanics and controls. The use of validating the survey results with in-game analytics and engagement metrics further strengthens the validity of the research being conducted.

\textbf{Reliability}: To help ensure reliability, the same setup, game environment and surveys will be used for all participants. Any and all instructions will be standardised, using text and/or video/audio
recordings to ensure that all participants are given the same instructions, without any bias that comes from the researcher. All data collection processes will be automated, further standardising the process
and ensuring reliability.

@@

\subsection{Ethical considerations}

@@

The small sample size of participants, along with the results show above indicate that the reliability of the results might be slightly skewed,
as due to having them replay the game multiple times, they were better able to learn the patterns of the FSM agents, especially on the 3rd playthrough,
which was the hard difficulty. This, along with the fact that its a single game, could limit the generalisability of the results,

\subsection{Summary of Findings}
From the analysis of the playtest data, we saw that on easy difficulty, FSM agents provided a more meaningful challenge, and a better player experience than RL agents.
This could have been due to players not having learned how to exploit the FSM agents yet, or the RL agent being too inconsistent at this difficulty.
As the difficulty increased, RL agents became more competitive, with the average score difference decreasing significantly, and player experience metrics such as engagement and immersion increasing.
Contrarily, FSMs became more predictable and easier to exploit, leading to a decrease in player engagement and immersion at higher difficulties.
This means that hypothesis is only partially supported, as while players reported higher levels of enjoyment and immersion when playing against RL agents, this was not the case at the easy difficulty.

@@

This study set out to compare the impact of Reinforcement Learning (RL) and Finite State Machine (FSM) agents on player experience in Real-Time Strategy (RTS) games.
The results from the playtest sessions indicate that in most cases, RL agents provide a more engaging and immersive experience for players.

The research questions were answered as follows: 
(1) RL agents generally outperformed FSMs in terms of player engagement at higher difficulty levels, while FSMs were preferred at easier levels; 
(2) The key factors influencing player experience between the two AI approaches were agent adaptability, predictability, and challenge;
(3) RL agents greatly improve player immersion, especially as difficulty increases, due to their adaptability and unpredictability, while FSMs suffered from predictability and were exploited by players;

As stated in the findings and discussion, the hypothesis was only partially supported, as players reported an improved experience when playing against RL agents
on all difficulties except for easy, where FSMs were preferred.

The limited number of participants and a more simplistic game design are the key limitations of this study, as the majority of participants were familiar with the game genre.
This likely influenced their ability to exploit the FSM agents, and made harder difficulties more desirable, as it better matched their skill level. This paired with having to replay the game multiple times on
harder difficulties may have skewed the results and their reliability. As for generalisability, due to it being a small scale study with a single simple game, the results may not be applicable to other RTS games,
or similar genres.

For future research, it is recommended to conduct a larger scale study with a more diverse player base, and to integrate Dynamic Difficulty Adjustment (DDA) systems to better balance the difficulty of the game.
This will help ensure that the game remains challenging across all skill levels, and remove the bias of having players repeat the game multiple times, increasing the difficulty each time.
Another recommendation is to explore the use of RL agents in other game genres, such as first-person shooters or action-adventure games, to see if the findings of this study can be generalised to other genres.
